---
title: "Assignment_5"
author: "Evan Z-G"
date: "2023-09-21"
output: pdf_document
---

```{r setup, include=FALSE}
suppressPackageStartupMessages({
  library(tidyverse, quietly = TRUE)   # loading ggplot2 and dplyr
})
options(tibble.width = Inf)   # Print all the columns of a tibble (data.frame)
```


Chapter 5, Problem 2
Create two scatter plots of gestational length and birth weight, one for each smoking status

```{r}
library(faraway)
data(phbirths)
#smoker <- phbirths %>%
#  group_by(smoke)
#smoke_yes <- filter(phbirths, smoke == TRUE)
#smoke_no <- filter(phbirths, smoke == FALSE)
ggplot(phbirths, aes(x=grams, y=gestate)) +
  geom_point() +
  facet_grid(cols = vars(smoke))
  
  
  #labs(x="Birth Weight (grams)", y="Gestational Length (weeks)")

```


Remove all the observations that are premature (less than 36 weeks). For the remainder of the problem, only use these full-term babies.

```{r}

mature <- phbirths %>% filter(., gestate > 35) 

ggplot(mature, aes(y=grams, x=gestate)) +
  geom_point() +
  facet_grid(cols = vars(smoke))
head(mature)
```


Fit the quadratic model

```{r}
model <- lm(grams ~ poly(gestate,2) * smoke, data=mature)
summary(model)
```


Add the model fitted values to the phbirths data frame along with the regression model confidence intervals.

```{r}

mature <- mature %>%
  dplyr::select( -matches('fit'), -matches('lwr'), -matches('upr') ) %>%
  cbind( predict(model, newdata=., interval='confidence') )

head(mature)


```


On your two scatter plots from part (a), add layers for the model fits and ribbon of uncertainty for the model fits.

```{r}
  ggplot(mature, aes(y=grams, x=gestate)) +
  geom_point() +
  facet_grid(cols = vars(smoke)) +
  geom_line( aes(y=fit) ) +
  geom_ribbon( aes( ymin=lwr, ymax=upr), alpha=.3 )
```

Create a column for the residuals in the phbirths data set using any of the following:

```{r}
mature$residuals = resid(model)

```


Create a histogram of the residuals.

```{r}
ggplot(mature, aes(x=residuals)) +
  geom_histogram() 
  
```

Chapter 6, Problem 2:


a.
```{r}
a <- 4
b <- 10
x <- runif(n=1, 0,10)  # one random value between 0 and 10 

if( x < a ){
  result <- 0     # Replace ???? with something appropriate!
}else if( x <= b ){
  result <- 1/(b-a)
}else{
  result <- 0
}
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))
```


b. i.
```{r}
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
if( (a<=x) & (x<=b) ){
  result <- 1/(b-a)
}else{
  result <- 0
}
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))
```


b. ii.
```{r}
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
if( (x<a) | (b<x) ){
  result <- 0
}else{
  result <- 1/(b-a)
}
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))

```
b. iii.
```{r}
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
result <- ifelse( a<=x & x<=b, 1/(b-a), 0 )
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))
```



3.a.
```{r}
library(ggplot2)
N <- 1000
df <- 2
x.grid <- seq(-3, 3, length=N)
data <- data.frame( 
  x = c(x.grid, x.grid),
  y = c(dnorm(x.grid), dt(x.grid, df)),
  type = c( rep('Normal',N), rep('T',N) ) )



for ( df in 2:30 ){
  # print out current value of df
  print( paste("In loop and df is now:", df) )  


  # make a nice graph
  myplot <- ggplot(data, aes(x=x, y=y, color=type, linetype=type)) +
  geom_line() +
  labs(title = paste('Std Normal vs t with', df, 'degrees of freedom'))

  # actually print the nice graph we made
  print(myplot)
  }
```

3. b.


```{r}
library(ggplot2)
N <- 1000
df <- 2
x.grid <- seq(-3, 3, length=N)
data <- data.frame( 
  x = c(x.grid, x.grid),
  y = c(dnorm(x.grid), dt(x.grid, df)),
  type = c( rep('Normal',N), rep('T',N) ) )


for( df in c(2,3,4,5,10,15,20,25,30) ){
  # print out current value of df
  print( paste("In loop and df is now:", df) )  

  # make a nice graph
  myplot <- ggplot(data, aes(x=x, y=y, color=type, linetype=type)) +
  geom_line() +
  labs(title = paste('Std Normal vs t with', df, 'degrees of freedom'))

  # actually print the nice graph we made
  print(myplot)
  }
```

4. a.

We can simulate rolling two 6-sided dice using the sample() function with the replace=TRUE option. Read the help file on sample() to see how to sample from the numbers 

```{r}
throw <- sample(1:6, size=2, replace = TRUE, prob = NULL)

throw
str(throw)
```
4. b.




```{r}
throws <- NULL
print(throws)
for( i in 1:24 ){
  throws[i] <- sample(1:6, size=2, replace = TRUE, prob = NULL) %>% sum()
print(throws[i])
game <- any( throws == 12 ) # Gives a TRUE/FALSE value
}
throws
print(game)
```


